---
title: "Lab 2: Feedback and Discussion"
author: "Nicky Wakim"
title-slide-attributes:
    data-background-color: "#C2352F"
date: "04/30/2025"
format: 
  revealjs:
    theme: "../simple_NW.scss"
    chalkboard: true
    slide-number: true
    show-slide-number: all
    width: 1955
    height: 1100
    footer: "Lab 2: Feedback and Discussion"
    html-math-method: mathjax
    highlight-style: ayu
execute:
  echo: true
  freeze: auto  # re-render only when source changes
---

```{r}
#| echo: false
#| warning: false
#| message: false

library(tidyverse)
library(gt)
library(janitor)
library(rstatix)
library(knitr)
library(gtsummary)
library(moderndive)
library(broom) 
library(here) 
library(ggplot2)
library(ggpubr)
library(boot) # for inv.logit()
library(lmtest)
library(epiDisplay)

```

## Notes from me and the TAs

- Using `ANY_HARDSHIP` or `MULT_HARDSHIPS`
  - If you look at the User Guide, this involves information from food security
- Using "Survey Questions"
  - Be very careful of this! 
- Interpreting estimated ORs
  - If your main variable had >6 categories I'm sorry if you interpreted every single one
- Make sure you are taking the inverse logit to find the predicted probability
- Make sure you are plotting the predicted probability, not the ORs
- The TAs left comments in the rubric! Please look at them! 

## Get into groups of 2-4

- No more than 4! 
- Try to group with people with whos you do not already work on labs 
- Introduce yourself if you do not know each other
- Share you html documents with each other (email, airdrop, etc.)

```{r}
#| echo: false
countdown::countdown(5)
```

## Introduce Research Question

- Share your research question with each other! 
  - Should be using the same outcome + your main variable
- Share your other variables

 

```{r}
#| echo: false
countdown::countdown(5)
```

## Double check code for variables

- Did everyone code their variables correctly?
- If you changed something to a factor, does it have the correct levels?

```{r}
#| echo: false
countdown::countdown(10)
```

## Discuss your confounders/effect modifiers

- Does not need to have a causal framework
- Do you think we need to adjust for certain variables?
- Do you think there will be interactions?
- If you think there are interactions, try to narrow it down to 1-2 variables


```{r}
#| echo: false
countdown::countdown(10)
```

## Contingency tables + bivariate analysis

- Based on your reasoning in the previous activity and the tables/plots, do you think we need to adjust for any variables?
- What trends are you seeing?
- Do you have any cells with low counts?

```{r}
#| echo: false
countdown::countdown(10)
```

## Simple logistic regression

- Does everyone have the correct interpretations of the estimated odds ratios?
- What's your reference group? 
- What trends are you seeing?

```{r}
#| echo: false
countdown::countdown(10)
```

## Predicted Probabilities

- Let's just check each other's code and probabilities
- What would your predicted probabilities look like if you had a confounder? effect modifier?

```{r}
#| echo: false
countdown::countdown(10)
```