---
title: "Lab 2: Feedback and Discussion"
author: "Nicky Wakim"
title-slide-attributes:
    data-background-color: "#C2352F"
date: "`r library(here); source(here('class_dates.R')); w5d2`"
format: 
  revealjs:
    theme: "../simple_NW.scss"
    chalkboard: true
    slide-number: true
    show-slide-number: all
    width: 1955
    height: 1100
    footer: "Lab 2: Feedback and Discussion"
    html-math-method: mathjax
    highlight-style: ayu
execute:
  echo: true
  freeze: auto  # re-render only when source changes
---

```{r}
#| echo: false
#| warning: false
#| message: false

library(tidyverse)
library(gt)
library(janitor)
library(rstatix)
library(knitr)
library(gtsummary)
library(moderndive)
library(broom) 
library(here) 
library(ggplot2)
library(ggpubr)
library(boot) # for inv.logit()
library(lmtest)
library(epiDisplay)

```

## Notes from me and the TAs




## Can I make this numeric?

- I've been getting this question for a few variables
- Some variables that might seem like we can convert to numeric 
  - PPEDUCAT
  - PPINCIMP
  - NUM_CHILD_HH
  - NUM_OWN_CHILD
  - NUM_FAM_SIZE

 

- To convert a categorical variable to numeric, we need to convince ourselves that the categorical variable has a linear relationship with the log-odds of food insecurity

- How do we do that?

## Check linearity!

-   When we treat a predictor as continuous, we need to make sure we have **linearity between the predictor and log-odds**
-   Easier to test this by fitting a logistic regression with the predictor

- Test linearity by fitting logistic regression with the predictor, then visually examining if predicted log-odds follows roughly linear pattern

 

- Let's test this out with income

## Set up the predicted log-odds

```{r}
#| echo: false

load("/Users/wakim/Library/CloudStorage/OneDrive-OregonHealth&ScienceUniversity/Projects/Teaching_projects/Teaching_DEIJ_Datasets/Datasets/Food_insecurity_2023/Data/Processed/wbns_data_2026-04-16.rda")
```

- First, I run a logistic regression model with `PPINCIMP` and `FOOD_INSEC`

```{r}
income_glm = wbns3 %>% glm(formula = FOOD_INSEC ~ PPINCIMP, family = binomial)
```

- Then I use `predict()` to calculate the log-odds for each category

```{r}
newdata = data.frame(PPINCIMP = levels(wbns3$PPINCIMP)) 
pred = predict(income_glm, newdata, se.fit=T, type = "link")

pred_link = data.frame(PPINCIMP = levels(wbns3$PPINCIMP), 
                       Pred = pred$fit, 
                       LL_CI1 = pred$fit - qnorm(1-0.05/2) * pred$se.fit, 
                       UL_CI1 = pred$fit + qnorm(1-0.05/2) * pred$se.fit) %>% 
  mutate(PPINCIMP = factor(PPINCIMP, levels = levels(wbns3$PPINCIMP)))
```

## Now I can plot the log-odds vs. income: what do we notice??

```{r}
#| code-fold: true

ggplot() + 
  labs(y = "Log-odds of Outcome") +
  geom_point(data = pred_link, aes(x = PPINCIMP, y=Pred), size=3) +
  geom_errorbar(data = pred_link, aes(x = PPINCIMP, y=Pred, ymin = LL_CI1, ymax = UL_CI1), width = 0.25) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## So I need to go back and map the levels to their midpoints {.smaller}

```{r}
wbns4 <- wbns3 %>%
  mutate(income_midpoint = case_when(
    PPINCIMP == "Less than $5,000"     ~ 2500,
    PPINCIMP == "$5,000 to $7,499"     ~ 6250,
    PPINCIMP == "$7,500 to $9,999"     ~ 8750,
    PPINCIMP == "$10,000 to $12,499"   ~ 11250,
    PPINCIMP == "$12,500 to $14,999"   ~ 13750,
    PPINCIMP == "$15,000 to $19,999"   ~ 17500,
    PPINCIMP == "$20,000 to $24,999"   ~ 22500,
    PPINCIMP == "$25,000 to $29,999"   ~ 27500,
    PPINCIMP == "$30,000 to $34,999"   ~ 32500,
    PPINCIMP == "$35,000 to $39,999"   ~ 37500,
    PPINCIMP == "$40,000 to $49,999"   ~ 45000,
    PPINCIMP == "$50,000 to $59,999"   ~ 55000,
    PPINCIMP == "$60,000 to $74,999"   ~ 67500,
    PPINCIMP == "$75,000 to $84,999"   ~ 80000,
    PPINCIMP == "$85,000 to $99,999"   ~ 92500,
    PPINCIMP == "$100,000 to $124,999" ~ 112500,
    PPINCIMP == "$125,000 to $149,999" ~ 137500,
    PPINCIMP == "$150,000 to $174,999" ~ 162500,
    PPINCIMP == "$175,000 to $199,999" ~ 187500,
    PPINCIMP == "$200,000 to $249,999" ~ 225000,
    PPINCIMP == "$250,000 or more"     ~ 300000
  ))
```

## Do same prediction, but now inclue midpoints!

```{r}
newdata = data.frame(PPINCIMP = levels(wbns4$PPINCIMP)) 
pred = predict(income_glm, newdata, se.fit=T, type = "link")

pred_link = data.frame(PPINCIMP = levels(wbns4$PPINCIMP), 
                       income_midpoint = sort(unique(wbns4$income_midpoint)), 
                       Pred = pred$fit, 
                       LL_CI1 = pred$fit - qnorm(1-0.05/2) * pred$se.fit, 
                       UL_CI1 = pred$fit + qnorm(1-0.05/2) * pred$se.fit) %>% 
  mutate(PPINCIMP = factor(PPINCIMP, levels = levels(wbns4$PPINCIMP)))
```

## And plot again with midpoints!

```{r}
#| code-fold: true

ggplot() + 
  labs(y = "Log-odds of Outcome") +
  geom_point(data = pred_link, aes(x = income_midpoint, y=Pred), size=3) +
  geom_errorbar(data = pred_link, aes(x = income_midpoint, y=Pred, ymin = LL_CI1, ymax = UL_CI1), width = 0.25) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Now to peer discussion

## Get into groups of 2-4ish

- Try to group with people with whos you do not already work on labs 
- Introduce yourself if you do not know each other
- Share you html documents with each other (email, airdrop, etc.)

```{r}
#| echo: false
countdown::countdown(5)
```

## Introduce Research Question

- Share your research question with each other! 
  - Should be using the same outcome + your main variable
- Share your other variables

 

```{r}
#| echo: false
countdown::countdown(5)
```


## Discuss your confounders/effect modifiers

- Does not need to have a causal framework
- Do you think we need to adjust for certain variables?
- Do you think there will be interactions?
- If you think there are interactions, try to narrow it down to 1-2 variables


```{r}
#| echo: false
countdown::countdown(10)
```

## Contingency tables + bivariate analysis

- Based on your reasoning in the previous activity and the tables/plots, do you think we need to adjust for any variables?
- What trends are you seeing?
- Do you have any cells with low counts?

```{r}
#| echo: false
countdown::countdown(10)
```

## Simple logistic regression

- Does everyone have the correct interpretations of the estimated odds ratios?
- What's your reference group? 
- What trends are you seeing?

```{r}
#| echo: false
countdown::countdown(10)
```

## Predicted Probabilities

- Let's just check each other's code and probabilities
- What would your predicted probabilities look like if you had a confounder? effect modifier?

```{r}
#| echo: false
countdown::countdown(10)
```