---
title: "Lab 1: Feedback and Discussion"
author: "Nicky Wakim"
title-slide-attributes:
    data-background-color: "#C2352F"
date: "`r library(here); source(here('class_dates.R')); w3d2`"
format: 
  revealjs:
    theme: "../simple_NW.scss"
    chalkboard: true
    slide-number: true
    show-slide-number: all
    width: 1955
    height: 1100
    footer: "Lab 1: Feedback and Discussion"
    html-math-method: mathjax
    highlight-style: ayu
execute:
  echo: true
  freeze: auto  # re-render only when source changes
---

```{r}
#| echo: false
#| warning: false
#| message: false

library(tidyverse)
library(gt)
library(janitor)
library(rstatix)
library(knitr)
library(gtsummary)
library(moderndive)
library(broom) 
library(here) 
library(ggplot2)
library(ggpubr)
library(boot) # for inv.logit()
library(lmtest)
library(epiDisplay)

```

## Mistake on my end

- I forgot to update the lab qmd for your work! So sorry!
- The TAs were lenient on the grading for this, but in the future make sure to complete the tasks from the instruction page!

## Notes from me and the TAs

- Using `ANY_HARDSHIP` or `MULT_HARDSHIPS`
  - If you look at the User Guide, this involves information from food security
  - We need to adapt the variable to exclude food security: use `case_when()` to construct it
  - Other option: use a different variable!!
- Using "Survey Questions": **please avoid using these**
  - Be very careful of this! 
- Activity 2.9: List any categorical variables that have less than 100 observations in a group
  - Not the number of observations for an entire variable, but check each level/group of the variable!
- Do not use multiple variables that capture the same information: **multicollinearity issues**
  - Example: `PPINCIMP` and `FPL50_100`
- Do NOT use variables that are measuring missing data unless you are interested in how missingness is related to the outcome
  - Example: `FAMSIZE_MISS`, `NUM_CHILD_HH_MISS`
- Make sure your image shows in your html
  - You need to have right file path!
  - How's the Rproj going?
  
## More notes: things that will lose points in Lab 2

- Purposefuly model selection
  - Dataset should have your outcome and 10 other variables
  - You should have a clear research question
- LASSO
  - You should have a dataset with a lot of variables (~51)
  - Start taking a look at the variables and think critically about potential multicollinearity issues
    - AND 
- Make sure your categorical variables are factors
  - Remake Table 1 in Lab 2 with the appropriate variable types


## LASSO: Updated variable selection

- There are a few more variables to remove

```{r}
#| eval: false

wbns2 = wbns %>% select(-starts_with("Q")) %>%  # remove survey questions
  select(-starts_with("V_")) %>%                # remove different versions of variables
  select(-ends_with("IMP")) %>%                 # remove imputed variables 
  select(-PPREG4, -NUM_SIBLING, -NUM_PARGUARD) %>%
  drop_na()
```

- This left me with a dataset with 51 variables and 6,658 observations


## Purposeful Model Selection: A note on the "Survey Questions"

- I said not to use the variables under this, but you can if you found one to be very interesting
- You just need to be extra cautious using these because they are not as nicely packaged/processed
- And some of them go into other variables!! 
  - Like `UNMETCARE_Y` already includes the survey questions about health care
- Some of the variable have A LOT of missing data

## For everyone

- If you had issues with removing the "(0)" and "(1)" from variable names WHILE preserving the factor class of the variables, here's some code that might help

```{r}
#| eval: false

wbns3 <- wbns2 %>%
  mutate(across(where(is.factor), ~ {
    levels(.) <- gsub("^\\(\\d+\\)\\s*", "", levels(.)) 
    .
  }))
```

## Get into groups of 2-4

- No more than 4! 
- Introduce yourself if you do not know each other
- Share you html documents with each other (email, airdrop, etc.)

```{r}
#| echo: false
countdown::countdown(5)
```

## Introduce your project type

- Are you doing LASSO or purposeful model selection?
- Share your research question with each other! 
- Share your other variables
- Is everyone using the appropriate variables?
- Did anyone see any potential issues with how many individuals are in each category of a variable?

 

- If time, did anyone want to use `ANY_HARDSHIP` or `MULT_HARDSHIPS`? How would you construct these variables?

```{r}
#| echo: false
countdown::countdown(15)
```

## Double check code and saved data

- Did everyone in the group successfully save the dataset?
- Can you open it right now?
- How many observations do you have after you dropped your NAs?

```{r}
#| echo: false
countdown::countdown(5)
```

## Compiled introductions

- What major points did you include in your introduction?
- Would you change anything after discussing with your group?

```{r}
#| echo: false
countdown::countdown(10)
```