---
title: "Lesson 13: Model Diagnostics"
author: "Nicky Wakim"
title-slide-attributes:
    data-background-color: "#C2352F"
date: "`r library(here); source(here('class_dates.R')); w6d2`"
format: 
  revealjs:
    theme: "../simple_NW.scss"
    chalkboard: true
    slide-number: true
    show-slide-number: all
    width: 1955
    height: 1100
    footer: "Lesson 13: Model Diagnostics"
    html-math-method: mathjax
    highlight-style: ayu
execute:
  echo: true
  freeze: auto 
editor_options: 
  chunk_output_type: console
---

```{r}
#| echo: false
library(tidyverse)
library(here)
library(aplore3)
library(regressinator)
library(ggplot2)
library(broom)
library(gt)
library(gtsummary)

theme_set(theme_minimal())

mean_age = mean(glow500$age) %>% round()
glow1 = glow500 %>% mutate(age_c = age - mean_age)
load(here("lessons", "12_Assessing_fit/data", "glow_results.Rda"))
# Assuming glow_m3 is the interaction model

```

# Learning Objectives

1. Understand the components of calculations for logistic regression diagnostics
2. Plot and determine observations where regression does not fit well or are influential using specific diagnostic values

## Review of model assessment so far

* Overall measurements of fit
* How well does the fitted logistic regression model predict the outcome?
* Different ways to measure the answer to this question



| Measure of fit | Hypothesis tested? | Equation | R code |
| --- | --- | --- | --- |
| Pearson residual | Yes | $X^2=\sum_{j=1}^{J}{r\left(Y_j,{\widehat{\pi}}_j\right)^2}$ | Not given |
| Hosmer-Lemeshow test | Yes | $\widehat{C}=\sum_{k=1}^{g}\frac{\left(o_k-n_k^\prime{\bar{\pi}}_k\right)^2}{n_k^\prime{\bar{\pi}}_k(1-{\bar{\pi}}_k)}$ | `hoslem.test()` |
| AUC-ROC | Kinda | Not given | `auc(observed, predicted)` |
| AIC | Only to compare models | $AIC = -2 \cdot \text{log-likelihood} + 2q$ | `AIC(model_name)` |
| BIC | Only to compare models | $BIC = -2 \cdot \text{log-likelihood} + q\text{log}(n)$ | `BIC(model_name)` |

## Today: Diagnostic Measures

* We move from overall fit to investigating **individual data** or **covariate pattern** data
* Make sure the overall measure has not been influenced by certain observations
* Assessed for **each covariate pattern** $j$ or each individual observation
* **Main diagnostic measurements we will look at:**
* Standardized Deviance residuals, $d_i$ (Outliers)
* Cook's distance, $D_i$ (Overall Influence)
* Change in coefficients, $DFBETA$ or $\Delta \widehat{\beta}_j$ (Coefficient Influence)
* Change in estimated probabilities, $DFFITS_i$ (Prediction Influence)



## Diagnostic Statistics Computation

* **Approach 1**: computed by covariate pattern (Hosmer-Lemeshow preference)
* **Approach 2**: individual observation approach
* Why consider both?
* When the number of covariate patterns is much smaller than $n$, we risk failing to identify poorly fit groups if we only look at individuals.
* Conversely, in models with continuous predictors, most individuals *are* their own covariate pattern.



## 1. Standardized Deviance Residuals ($d_i$)

* Identifies **outliers**: observations where the model's prediction is far from the truth.
* Standardized to account for leverage ($h_i$):

$$d_i = \frac{\text{deviance residual}_i}{\sqrt{1-h_i}}$$


* **Rule of Thumb:** Large values (typically $|d_i| > 3$) indicate a covariate pattern that the model does not fit well.

## 2. Cook's Distance ($D_i$)

* Measures the **aggregate influence** of a covariate pattern on all regression coefficients.
* It asks: "How much would the whole model change if we deleted this pattern?"
* Combined measure of leverage ($h_i$) and the residual.
* **Rule of Thumb:** Flag values where $D_i > 1$ or $D_i > 4/n$.

## 3. Change in Coefficients (DFBETA)

* Measures the change in a **specific** coefficient $\widehat{\beta}_j$ if an observation is excluded.
* $$\Delta \widehat{\beta}_j = \frac{\widehat{\beta}_j - \widehat{\beta}_{j(-i)}}{SE(\widehat{\beta}_{j(-i)})}$$


* Useful for ensuring our variable of interest (e.g., Prior Fracture) isn't being driven by a single weird data point.
* **Rule of Thumb:** Look for $|DFBETA| > 2/\sqrt{n}$.

## 4. Change in Probabilities (DFFITS)

* Measures the impact on the **predicted probability** (the "fit").
* It is the number of standard deviations the fitted value changes when an observation is removed.
* **Rule of Thumb:** Flag if $|DFFITS| > 2\sqrt{p/n}$.

# Visual Assessment

::: lob
2. Plot and determine observations where regression does not fit well or are influential using specific diagnostic values
:::

## How do we get these values in R?

* We use `broom::augment` to get the core residuals and `influence.measures` for the rest.

```{r}
# Augment for residuals
aug_glow = augment(glow_m3, type.predict = "response", type.residuals = "deviance")

# Calculate influence measures (Cook's D, DFFITS, DFBETAs)
infl = influence.measures(glow_m3)
glow_dx = cbind(aug_glow, as.data.frame(infl$infmat)) %>%
          rename(std.dev.res = .std.resid, cooksd = cook.d)

```

## GLOW: Standardized Deviance Residuals

::: columns
::: column

* Look for $|d_i| > 3$.
* Points at the top/bottom indicate patterns where $Y$ is the opposite of what we expected.
:::
::: column

```{r fig.height=7, fig.width=8}
#| code-fold: true
ggplot(glow_dx, aes(x = .fitted, y = std.dev.res)) +
  geom_point() +
  geom_hline(yintercept = c(-3, 3), col = "red", linetype="dashed") +
  labs(x = "Predicted Probability", y = "Std. Deviance Residual")

```

:::
:::

## GLOW: Cook's Distance

::: columns
::: column

* Threshold: $4/500 = 0.008$.
* Points above the line are "globally" influential on our results.
:::
::: column

```{r fig.height=7, fig.width=8}
#| code-fold: true
ggplot(glow_dx, aes(x = .fitted, y = cooksd)) +
  geom_point() +
  geom_hline(yintercept = 0.008, col = "red", linetype="dashed") +
  labs(x = "Predicted Probability", y = "Cook's Distance")

```

:::
:::

## Summary & Reporting

**Methods:** We assessed model fit using Standardized Deviance residuals (outliers), Cook’s Distance (overall influence), and DFBETAs (coefficient stability).

**Results:** AUC-ROC was 0.68. We identified $X$ observations with Cook's distance $> 0.008$. Sensitivity analysis showed coefficient estimates remained stable; thus, all observations were retained.

**Discussion:** High influence points were investigated and found to represent valid clinical scenarios (e.g., older patients with prior fractures who did not refracture), and were kept in the model.
