---
title: "Lesson 13: Model Diagnostics"
author: "Nicky Wakim"
title-slide-attributes:
    data-background-color: "#C2352F"
date: "`r library(here); source(here('class_dates.R')); w7d1`"
format: 
  revealjs:
    theme: "../simple_NW.scss"
    chalkboard: true
    slide-number: true
    show-slide-number: all
    width: 1955
    height: 1100
    footer: "Lesson 13: Model Diagnostics"
    html-math-method: mathjax
    highlight-style: ayu
execute:
  echo: true
  freeze: auto  # re-render only when source changes
editor_options: 
  chunk_output_type: console
---

```{r}
#| echo: false

library(tidyverse)
#library(textrecipes)
library(here)
library(aplore3)
library(regressinator)
library(ggplot2)
library(broom)
library(gt)
library(gtsummary)


theme_set(theme_minimal())


mean_age = mean(glow500$age) %>% round()
glow1 = glow500 %>% mutate(age_c = age - mean_age)
load(here("lessons", "12_Assessing_fit/data", "glow_results.Rda"))
source(here("lessons", "13_Model_diagnostics", "Logistic_Dx_Functions.R"))

# glow_m3 = glm(fracture ~ priorfrac + armassist, data = glow1, family = binomial)

theme_set(theme_bw(base_size = 22))
```


# Learning Objectives

1. Understand the components of calculations for logistic regression diagnostics

2. Plot and determine observations where regression does not fit well or are influential using specific diagnostic values

```{css, echo=FALSE}
.reveal code {
  max-height: 80% !important;
}
```


## Review of model assessment so far

-   Overall measurements of fit

    -   How well does the fitted logistic regression model predict the outcome?
    
    -   Different ways to measure the answer to this question
    
| Measure of fit       | Hypothesis tested?     | Equation                                                                                                            | R code                     |
|----------------|----------------|------------------------|----------------|
| Pearson residual     | Yes                    | $X^2=\sum_{j=1}^{J}{r\left(Y_j,{\widehat{\pi}}_j\right)^2}$                                                             | Not given                  |
| Hosmer-Lemeshow test | Yes                    | $\widehat{C}=\sum_{k=1}^{g}\frac{\left(o_k-n_k^\prime{\bar{\pi}}_k\right)^2}{n_k^\prime{\bar{\pi}}_k(1-{\bar{\pi}}_k)}$ | `hoslem.test()`            |
| AUC-ROC              | Kinda                  | Not given                                                                                                           | `auc(observed, predicted)` |
| AIC                  | Only to compare models | $AIC = -2 \cdot \text{log-likelihood} + 2q$                                                                         | `AIC(model_name)`          |
| BIC                  | Only to compare models | $BIC = -2 \cdot \text{log-likelihood} + q\text{log}(n)$                                                             | `BIC(model_name)`          |

## Next week for model assessment


-   Numerical problems

    -   Assess pre and post model fit
    -   Numerical problems often depend on the final model (which variables and interactions are included)

 

-   Different numerical problems to look out for

    -   Zero cell count
    -   Complete separation
    -   Multicollinearity

 

-   **Today:** We now use model diagnostics to identify any **observations** that the model does not fit well

    
## Assumptions: Linear Regression vs. Logistic Regression

::: columns
::: column
::: axiom
::: axiom-title
Linear Regression Assumptions
:::
::: axiom-cont
- Linearity of relationship between **$Y$** and covariates
- Independence of observations
- Normality of $Y$ given $X$ (residuals)
- Equality of variance of residuals (homoscedasticity)
:::
:::
:::
::: column

::: proof1
::: proof-title
Logistic Regression Assumptions
:::
::: proof-cont
- Linearity of relationship between **logit of outcome** and covariates
- Independence of observations
- Distribution of $Y$ given $X$ follows the prescribed one (binomial in our case)
:::
:::

:::
:::

- Back in [Lesson 5: Simple Logistic Regression](https://nwakim.github.io/BSTA_513_S25/lessons/05_Simple_logistic_reg/05_Simple_logistic_reg.html#/title-slide), we walk through linear regression assumptions that do not hold for logistic regression
- Because our assumptions for logistic regression are a little different than linear regression, we **need slightly different diagnostic tools** to assess observations that may be outliers or influential points

# Learning Objectives

::: lob
1. Understand the components of calculations for logistic regression diagnostics
:::

2. Plot and determine observations where regression does not fit well or are influential using specific diagnostic values

## Review of Number of Covariate Patterns {visibility="hidden"}

-   Covariate patterns are the **unique covariate combinations** that are observed

 

-   **For example:** model contains two binary covariates (history of fracture and smoking status), there will be **4 unique combination of these factors**

    -   This model has **4 covariate patterns**
    -   Subjects can be divided into 4 groups based on the covariates’ values
 
 
   
-   When we have **continuous covariates**, the number of covariate patterns will be close to the **number of individuals in the dataset**

## From overall fit measure to diagnostic measures {visibility="hidden"}

-   Now we need to investigate diagnostics looking at *individual* data or *covariate pattern* data

    -   Make sure the overall measure has not been influenced by certain observations

 


-   Assessed for **each covariate pattern** $j$ or **each individual observation**

 

- **Main diagnostic measurements we will look at: **
  - Standardized Deviance residuals, $d_i$ (outliers)
  - Cook's distance, $D_i$
  - Change in coefficients, $DFBETA$ or $\Delta \widehat{\beta}_j$
  - Change in estimated probabilities, $DFFITS_i$

## Diagnostic Statistics Computation {visibility="hidden"}

-   Two diagnostic statistics computation approach

    -   **Approach 1**: computed by covariate pattern

        -   Recommendation of Hosmer-Lemeshow textbook
        -   Identify outliers as group that shares the same covariate values (in the same covariate pattern)
        -   Can be better if you have very few covariate patterns
    
     
    
    -   **Approach 2**: individual observation approach

        -   SAS and R use this approach automatically
        -   Identify outliers as individual

 

-   Why prefer covariate patterns approach?

    -   When the number of covariate pattern is much smaller than $n$, there is risk that we may fail to identify influential and/or poorly fit covariate patterns using individual based on residual
    
## Poll Everywhere Question 1 {visibility="hidden"}

## Diagnostic Statistics Computation (2/2) {visibility="hidden"}

Consider a covariate pattern with $m_j$ subjects with estimated probability $\widehat\pi_j$

-   Pearson residual computed by individual $$r_i=-\sqrt{\frac{{\widehat{\pi}}_j}{(1-{\widehat{\pi}}_j)}}$$

-   Pearson residual computed by covariate pattern $$r_j=-\sqrt{m_j}\sqrt{\frac{{\widehat{\pi}}_j}{(1-{\widehat{\pi}}_j)}}$$

-   **Difference between above residuals will be large if $m_j$ is large**: usually a problem if less covariate patterns

    -   Residual from covariate pattern will identify poorly fit covariate patterns



## (Some) Diagnostics of Logistic Regression 

-   Measures at the observation level for model diagnostics!

-   Model diagnostics of logistic regression can be assessed by checking for **outliers**:

      - (1) Standardized Deviance residuals, $d_i$

 

-   Model diagnostics of logistic regression can be assessed by checking how **influential** an observation is:

      - (2) Cook's distance, $D_i$
      - (3) Change in coefficients, $DFBETA$ or $\Delta \widehat{\beta}_j$
      - (4) Change in estimated probabilities, $DFFITS_i$

## 1. Standardized Deviance Residuals ($d_i$)

* Identifies **outliers**: observations where the **model's prediction is far from the truth**.

$$\text{deviance residual}_i = \text{sign}(y_i - \widehat{\pi}_i) \cdot \sqrt{-2\left[y_i\log\left(\dfrac{y_i}{\widehat{\pi}_i}\right) + (1-y_i)\log\left(\dfrac{1-y_i}{1-\widehat{\pi}_i}\right)\right]}$$

* Standardized to account for leverage ($h_i$, measure of distance between observed predictors and mean predictors):

$$d_i = \frac{\text{deviance residual}_i}{\sqrt{1-h_i}}$$



* **Rule of Thumb:** Large values (typically $|d_i| > 3$) indicate that the model does not fit observation $i$ well

## 2. Cook's Distance ($D_i$)

* Measures the **aggregate influence** of an observation on all regression coefficients.
* It asks: "How much would the whole model change if we deleted this pattern?"
* Combined measure of leverage ($h_i$) and the residual, with $k$ coefficients.

$$D_i = \dfrac{d_i^2 h_i}{(k-1)(1-h_i)^2}$$

* **Rule of Thumb:** Flag values where $D_i > 1$ or $D_i > 4/n$

## 3. Change in Coefficients (DFBETA)

* Measures the change in a **specific** coefficient $\widehat{\beta}_j$ if an observation is excluded.

$$\Delta \widehat{\beta}_j = \frac{\widehat{\beta}_j - \widehat{\beta}_{j(-i)}}{SE(\widehat{\beta}_{j(-i)})}$$


* Useful for ensuring our variable of interest (e.g., Prior Fracture) isn't being driven by a single observation
- This is different than the change in coefficient when we remove a predictor from the model (e.g., change in coefficient for prior fracture when we remove age from the model)
* **Rule of Thumb:** Look for $|DFBETA| > \dfrac{2}{\sqrt{n}}$.

## 4. Change in fitted probabilities (DFFITS)

* Measures the impact on the **predicted probability** (the "fit")
* It is the number of standard deviations the fitted value changes when an observation is removed


* **Rule of Thumb:** Flag if $|DFFITS| > 2\sqrt{k/n}$


# Learning Objectives

1. Understand the components of calculations for logistic regression diagnostics

::: lob
2. Plot and determine observations where regression does not fit well or are influential using specific diagnostic values
:::

## Recall the model we fit: GLOW Study with interactions

-   **Outcome variable:** any fracture in the first year of follow up (FRACTURE: 0 or 1)

-   **Risk factor/variable of interest:** history of prior fracture (PRIORFRAC: 0 or 1)

-   **Potential confounder or effect modifier:** age (AGE, a continuous variable)

-   Fitted model with interactions: $$\begin{aligned} \text{logit}\left(\widehat\pi(\mathbf{X})\right) & = \widehat\beta_0 &+ &\widehat\beta_1\cdot I(\text{PF}) & + &\widehat\beta_2\cdot Age& + &\widehat\beta_3 \cdot I(\text{PF}) \cdot Age \\  
    \text{logit}\left(\widehat\pi(\mathbf{X})\right) & = -1.376 &+ &1.002\cdot I(\text{PF})& + &0.063\cdot Age& -&0.057 \cdot I(\text{PF}) \cdot Age
    \end{aligned}$$

 
    
-   Lesson 12: determined the overall fit of this model
-   Today: determine the if any observations/covariate patterns that model does not fit well

## Recall our friend `augment()`?

```{r}
# Augment for residuals
library(broom)
aug_glow = augment(glow_m3, type.predict = "response", type.residuals = "deviance")

# Calculate influence measures (Cook's D, DFFITS, DFBETAs)
infl = influence.measures(glow_m3)
dx_glow = cbind(aug_glow, as.data.frame(infl$infmat)) %>%
          dplyr::select(-c(cook.d, cov.r, hat))
```

## Key to the values

::: columns
::: {.column width="48%"}
- `.fitted`: Estimated probability of events ($\widehat{\pi}$)
- `.resid`: Deviance residual
- `.hat`: Leverage
- `.sigma`: Standard error of the residuals
- `.cooksd`: Cook's distance 
- `.std.resid`: Standardized deviance residual
- `dfb.1_`: Change in coef. estimate for intercept
- `dfb.prrY`: Change in coef. estimate for prior fracture
- `dfb.age_c`: Change in coef. estimate for age
- `dfb.pY:_`: Change in coef. estimate for interaction between age and prior fracture
- `dffit`: Change in predicted probability of fracture
:::

::: {.column width="52%"}
```{r}
glimpse(dx_glow, width=50)
```
:::
:::

## Cutoffs and visual assessment for diagnostics

-   The plots allow us to identify those covariate patterns that are...

    -   Poorly fit 

        -   Large values of standardized deviance residuals ($d_i$)

    -   Influential on estimated coefficients 

        -   Large values of Cook's distance ($D_i$), change in coefficient estimates ($\Delta \widehat{\beta}_j$), and change in predicted probabilities ($DFFITS_i$)
        
-   We plot the diagnostic values **against the predicted probabilities** to see if there are any patterns in the poorly fit or influential observations

## GLOW Study: Standardized Deviance Residuals

::: columns
::: column

- Look for $|d_i| > 3$
- Points at the top/bottom indicate patterns where $Y$ is the opposite of what we expected

```{r}
dx_glow %>% filter(abs(.std.resid) > 3) |> 
  dplyr::select(priorfrac, age_c, 
                .fitted, .std.resid)
```
:::

::: column

```{r fig.height=7, fig.width=9}
#| code-fold: true
ggplot(dx_glow, aes(x = .fitted, y = .std.resid)) +
  geom_point() +
  geom_hline(yintercept = c(-3, 3), col = "red", linetype="dashed") +
  labs(x = "Predicted Probability", y = "Std. Deviance Residual")
```

:::
:::

## GLOW Study: Cook's Distance

::: columns
::: column

* Threshold: $D_i > 1$
* Points above the line are influential on our results.

```{r}
dx_glow %>% filter(.cooksd > 1) %>%
  dplyr::select(priorfrac, age_c, .fitted, .cooksd)
```

:::
::: column

```{r fig.height=7, fig.width=8}
#| code-fold: true
ggplot(dx_glow, aes(x = .fitted, y = .cooksd)) +
  geom_point() +
  geom_hline(yintercept = 1, col = "red", linetype="dashed") +
  labs(x = "Predicted Probability", y = "Cook's Distance")
```

:::
:::

## GLOW Study: Change in coefficient estimates (for prior fracture)

::: columns
::: column
-   Book recommends flagging observsations if change in coefficient estimates are greater than $2/\sqrt{n}$ 


```{r}
dx_glow %>% filter(dfb.prrY > 2/sqrt(500)) %>% 
  select(priorfrac, age_c, .fitted, dfb.prrY) %>%
  head(15)
```

:::

::: column
```{r fig.height=7, fig.width=8}
#| code-fold: true
#| code-summary: "To make the plot"

ggplot(dx_glow) + 
  geom_hline(yintercept = 2/sqrt(500), col = "red", linetype="dashed", size=2) +
  geom_point(aes(x=.fitted, y=dfb.prrY), size = 3) + 
  xlab("Estimated/Predicted Probability of Fracture") +
  ylab("Change in Coefficient Estimates") +
  theme(text = element_text(size = 26)) + xlim(0,1)
```
:::
:::

## GLOW Study: Changed in predicted probability

::: columns
::: column
- Flag if $|DFFITS| > 2\sqrt{k/n}$


```{r}
dx_glow %>% filter(abs(dffit) > 2*sqrt(4/500)) %>% 
  select(priorfrac, age_c, .fitted, dffit) %>%
  head(15)
```

:::

::: column
```{r fig.height=7, fig.width=8}
#| code-fold: true
#| code-summary: "To make the plot"

ggplot(dx_glow) + 
  geom_hline(yintercept = 2*sqrt(4/500), col = "red", linetype="dashed", size=2) +
  geom_point(aes(x=.fitted, y=dffit), size = 3) + 
  xlab("Predicted Probability of Fracture") +
  ylab("Change in Predicted Probability") +
  theme(text = element_text(size = 26)) + xlim(0,1)
```
:::
:::

## After identifying points

-   Do a data quality check

    -   Unless you have a very good reason to believe the data are not measured correctly, then we leave it in
    -   Recall our [slide from Linear Models (BSTA 512/612)](https://nwakim.github.io/BSTA_512_W25/lessons/15_MLR_Diagnostics/15_MLR_Diagnostics.html?q=outlier#/how-do-we-deal-with-influential-points) 
    -   Common to do nothing
    
 

-   We are not too worried about a few outliers
    
 

-   If we have quite a lot of outliers that are highly influential, potential reasons can be considered:

    -   The link used in logistic regression model is not appropriate for outcome

        -   This is usually unlikely, since logistic regression model is very flexible (think back to why we transformed our outcome from binary form)

    -   One or more important covariates missing in the model

        -   At least one of the covariates in the model has been entered in the wrong scale (think age-squared vs. age)
        
        
## How would I report this? (Combining all model assessment)

-   Assuming I have not checked other final models (no other models to compare AIC/BIC or AUC with)

**Methods:** Our final logistic regression model consisted of the outcome, fracture, and predictors including prior fracture, age, and their interaction. To assess the overall model fit, we calculated the AUC-ROC. We identified observations with high standardized deviance residual (absolute value greater than 3), Cook's distance (greater than 1), and change in predicted probability (greater than 0.179). No identified observations were omitted.

 

**Results:** The AUC-ROC was 0.68. Out of 500 observations, we identified 26 with high change in predicted probability. 

 

**Discussion:** 

-   AUC-ROC low: Included covariates were pre-determined 

-   Influential points were kept in because all observations were within feasible range of the predictors and outcome. (we could try age-squared and see if that helps AUC and/or diagnostics)



