---
title: "Lab 4: Feedback and Discussion"
author: "Nicky Wakim"
title-slide-attributes:
    data-background-color: "#C2352F"
date: "`r library(here); source(here('class_dates.R')); w9d2`"
format: 
  revealjs:
    theme: "../simple_NW.scss"
    chalkboard: true
    slide-number: true
    show-slide-number: all
    width: 1955
    height: 1100
    footer: "Lab 2: Feedback and Discussion"
    html-math-method: mathjax
    highlight-style: ayu
execute:
  echo: true
  freeze: auto  # re-render only when source changes
---

```{r}
#| echo: false
#| warning: false
#| message: false

library(tidyverse)
library(gt)
library(janitor)
library(rstatix)
library(knitr)
library(gtsummary)
library(moderndive)
library(broom) 
library(here) 
library(ggplot2)
library(ggpubr)
library(boot) # for inv.logit()
library(lmtest)
library(epiDisplay)

```

## Get into groups of 2-4

- No more than 4! 
- Get into groups that did the **same model selection strategy**
- Introduce yourself if you do not know each other
- Share you html documents with each other (email, airdrop, etc.)

```{r}
#| echo: false
countdown::countdown(5)
```

## Did you go through your model selection strategy correctly?

- For purposeful model selection:
  - Did you follow the steps from 512/612 correctly?
  - Did you have any questions about the steps?
  - Did you start with the 10 variables?

- For LASSO:
  - Did you have any overarching questions?
  - Did you remove the problem variables?
  - Did you use cross-validation to select lambda?
  - Did you assess your model with your testing set with AUC?
  - Did you omit correlated variables?

```{r}
#| echo: false
countdown::countdown(25)
```

## Draft your methods and results section!

- If you did purposeful model selection, 
  - Take a look at your methods section from BSTA 512/612!! Make updates from my comments
- If you did LASSO, 
  - Brainstorm with each other on what should be included
  - Your process should be nearly identical
- From yuor final model, what are the important results to report?

```{r}
#| echo: false
countdown::countdown(15)
```

