---
title: "Lesson 14: Model Building V2"
subtitle: "With an emphasis on prediction"
author: "Nicky Wakim"
title-slide-attributes:
    data-background-color: "#C2352F"
date: "05/14/2025"
format: 
  revealjs:
    theme: "../simple_NW.scss"
    chalkboard: true
    slide-number: true
    show-slide-number: all
    width: 1955
    height: 1100
    footer: "Lesson 14: Model Building"
    html-math-method: mathjax
    highlight-style: ayu
execute:
  echo: true
  freeze: auto  # re-render only when source changes
editor_options: 
  chunk_output_type: console
---

```{r}
#| echo: false
#| warning: false
#| message: false

library(gt)
library(gtsummary)
library(tidyverse)
library(glmnet) # The standard engine for LASSO
library(pROC)   # For evaluation
library(here)
library(aplore3)
library(openintro)
library(tidyr)

theme_set(theme_minimal())

mean_age = mean(glow500$age) %>% round()
glow1 = glow500 %>% mutate(age_c = age - mean_age)

set.seed(513613)
```

# Learning Objectives

1. Understand the place of LASSO regression within association and prediction modeling for binary outcomes.

2. **Understand the standard workflow for the `glmnet` package.**

3. Understand how penalized regression is a form of model/variable selection.

4. Perform LASSO regression using `glmnet` and find the optimal penalty ($\lambda$).

# Learning Objectives

::: lob
1. Understand the place of LASSO regression within association and prediction modeling for binary outcomes.
:::

## Bias-variance trade off

- Recall from 512/612: MSE can be written as a function of the bias and variance
  $$MSE = \text{bias}\big(\widehat\beta\big)^2 + \text{variance}\big(\widehat\beta\big)$$

- **For the same data:**
    - **More covariates:** less bias, more variance (Overfitting)
    - **Fewer covariates:** more bias, less variance (Underfitting)



## The goals of association vs. prediction

::::::::::: columns
:::::: column
::::: definition
::: def-title
Association / Explanatory
:::
- **Goal:** Understand a specific variable's effect after adjusting for others.
- Focus on odds ratios and p-values.
:::::
::::::

:::::: column
::::: proposition
::: prop-title
Prediction
:::
- **Goal:** Most precise prediction of the response variable.
- Interpreting coefficients is secondary.
- **Shrinkage methods** (like LASSO) help by removing noise.
:::::
::::::
:::::::::::

# Learning Objectives

::: lob
2. Understand the standard workflow for the `glmnet` package.
:::

## Why `glmnet`?

- It is the "industry standard" for penalized regression.
- It is computationally efficient and fast.
- **Key Difference:** It does not use the `y ~ x1 + x2` formula syntax. It requires:
    1. A **numeric matrix** of predictors ($X$).
    2. A **numeric vector** for the outcome ($Y$).

## Preparing the Data: Matrix format

Since `glmnet` needs a matrix, we use `model.matrix()` to handle our categorical variables automatically.

```{r}
# Define the outcome
y <- glow1$fracture

# Define the predictor matrix (removing intercept and ID columns)
# The [-1] removes the intercept column created by model.matrix
x <- model.matrix(fracture ~ priorfrac + age_c + premeno + momfrac + armassist + smoke, 
                  data = glow1)[, -1]

head(x)
```

# Learning Objectives

::: lob
3. Understand how penalized regression is a form of model/variable selection.
:::

## Three types of penalized regression

The `alpha` parameter in `glmnet` determines the method:

- **Lasso (`alpha = 1`):** Shrinks coefficients to exactly zero. Great for variable selection.
- **Ridge (`alpha = 0`):** Shrinks coefficients toward zero, but never exactly zero.
- **Elastic Net (`0 < alpha < 1`):** A weighted average of both.

# Learning Objectives

::: lob
4. Perform LASSO regression using `glmnet` and find the optimal penalty ($\lambda$).
:::

## Step 1: Splitting data

We split the matrix and vector into training and testing sets.

```{r}
train_rows <- sample(1:nrow(x), 0.8 * nrow(x))

x_train <- x[train_rows, ]
x_test  <- x[-train_rows, ]

y_train <- y[train_rows]
y_test  <- y[-train_rows]
```

## Step 2: Fit LASSO and Tune $\lambda$

Instead of picking a random penalty, we use **Cross-Validation** (`cv.glmnet`) to find the best $\lambda$.

```{r}
# family = "binomial" for logistic regression
# alpha = 1 for LASSO
cv_fit <- cv.glmnet(x_train, y_train, family = "binomial", alpha = 1, 
                    thresh = 1e-12)

y = as.numeric(y_train)-1
x = x_train
beta = coef(cv_fit, s = "lambda.min", exact = TRUE, x = x, y = y)
lambda = cv_fit$lambda.min

# Example for Lasso inference
library(selectiveInference)
# You need the original x, y, and the fixed lambda
out = fixedLassoInf(x, y, beta, lambda, family="binomial")
out

# 2. Transform the output into a "tidy" tibble
lasso_inference_df <- tibble(
  term      = names(out$vars),
  estimate  = exp(out$coef0),
  std.error = out$sd,
  statistic = out$z,
  p.value   = out$pv,
  conf.low  = exp(out$ci[,1]),
  conf.high = exp(out$ci[,2])
)

glm_mod = glow1[train_rows, ] %>%
  glm(formula = fracture ~ priorfrac + age_c + premeno + momfrac + armassist + smoke,
      family = "binomial")


tidy_df <- tidy(glm_mod, conf.int = T, exponentiate = T)

lasso_to_plot <- tibble(
  term      = colnames(x)[out$vars], 
  estimate  = exp(out$coef0),
  std.error = out$sd,
  statistic = out$z,
  p.value   = out$pv,
  conf.low  = exp(out$ci[,1]),
  conf.high = exp(out$ci[,2]),
  model     = "LASSO (Selective Inference)"
)

tbl_regression(lasso_inference_df)

# 2. Clean up and label the GLM data
glm_to_plot <- tidy_df %>%
  filter(term != "(Intercept)") %>%
  mutate(model = "Standard GLM (Logistic)") %>%
  # Ensure columns match for binding
  dplyr::select(term, estimate, std.error, statistic, p.value, conf.low, conf.high, model)

# 3. Combine them
combined_data <- bind_rows(lasso_to_plot, glm_to_plot)

# 4. Create the Forest Plot
ggplot(combined_data, aes(y = term, x = estimate, color = model)) +
  # Add vertical null line
  geom_vline(xintercept = 1, linetype = "dashed", color = "black", alpha = 0.5) +
  # Use position_dodge to prevent the points/lines from overlapping
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), 
                 height = 0.3, 
                 position = position_dodge(width = 0.5)) +
  geom_point(size = 3, position = position_dodge(width = 0.5)) +
  # Use Log scale for x-axis to balance ORs correctly
  scale_x_log10() +
  # Custom colors
  scale_color_manual(values = c("LASSO (Selective Inference)" = "#C2352F", 
                                "Standard GLM (Logistic)" = "#2F5597")) +
  labs(
    title = "Comparison of Model Estimates",
    subtitle = "Odds Ratios and 95% Confidence Intervals",
    x = "Odds Ratio (Log Scale)",
    y = NULL,
    color = "Model Type"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(face = "bold"),
    legend.position = "bottom"
  )
```

```{r}
tidy_df <- tidy(glm_mod, conf.int = T, exponentiate = T) %>%
  filter(term != "(Intercept)")

ggplot(tidy_df, aes(y = term, x = estimate)) +
  # Add a vertical line at the null (OR = 1)
  geom_vline(xintercept = 1, linetype = "dashed", color = "red") +
  # Add the horizontal confidence intervals
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  # Add the point estimates
  geom_point(size = 3, color = "#C2352F") +
  theme_minimal() +
  theme(axis.text.y = element_text(face = "bold"))
```



## Step 3: Identify Selected Variables

We can view the coefficients at the "best" lambda (the one that minimizes deviance).

```{r}
# Variables with a "." were dropped by the model
coef(cv_fit, s = "lambda.min")
```

## Step 4: Prediction and Evaluation

Finally, we test our model's performance on the testing set.

```{r}
# Get predicted probabilities
preds <- predict(cv_fit, newx = x_test, s = "lambda.min", type = "response")

# Calculate AUC
lasso_roc <- roc(y_test, as.vector(preds))
lasso_roc$auc
```

## Step 5: Visualizing the ROC Curve

```{r}
#| fig-align: center
plot(lasso_roc, col = "#C2352F", lwd = 3, main = "LASSO Performance (Testing Set)")
```

## Summary

- **`glmnet`** requires data as a matrix (`x`) and a vector (`y`).
- We use **`cv.glmnet`** to automatically find the best penalty ($\lambda$).
- LASSO performing **variable selection** by forcing some coefficients to zero.
- Prediction accuracy is assessed on a held-out **testing set**.
