---
title: "Lesson 1: Welcome!"
author: "Nicky Wakim"
title-slide-attributes:
    data-background-color: "#C2352F"
date: "`r library(here); source(here('class_dates.R')); w1d1`"
format: 
  revealjs:
    theme: "../simple_NW.scss"
    chalkboard: true
    slide-number: true
    show-slide-number: all
    width: 1955
    height: 1100
    footer: "Lesson 1: Welcome"
    html-math-method: mathjax
    highlight-style: ayu
execute:
  echo: true
  freeze: auto  # re-render only when source changes
---

```{r setup, include=FALSE}
library(here)
source(here("class_dates.R"))
```

# Still me!

## Course Learning Objectives

At the end of this course, students should be able to...

1.  Apply and interpret some hypothesis-testing procedures for two-way and three-way **contingency tables**

2.  Compute and **interpret measures of association** for **binary and ordinal data**.

3.  Calculate and correctly interpret **odds ratios using logistic regression**, make comparison across groups and examine relationship between binary outcome and predictor variables.

4.  Apply appropriate **model-building strategies for logistic regression**. Effectively use statistical computing packages for contingency table and logistic regression procedures.

5.  Perform **Poisson regression** analysis using count data and interpret model estimates, make comparison across groups and examine relationship between outcome and predictor variables.

6.  Coherently **summarize methods and results of data analyses**, and discuss in context of original health-related research questions to audiences with varied statistical background.

## New way to record/view lectures

-   I will be using Echo360 to automatically record time from 1-3pm in this room!

-   It will be LIVE

    -   So you can watch in real time, but won't be able to interact with us

    -   Might have a 10 second lag

    -   No need to tell me if you cannot make class

-   [Here is the link to the Echo site!](https://echo360.org/section/87bbad0a-58f2-4758-a290-b95378fbbab4/home) I'll keep posting it on the weekly pages

    -   Everyone should be added! Please check that you have access.

-   I don't have to post anything after class, so no issues with me forgetting to post it

## Some important tasks

-   Star the class website: <https://nwakim.github.io/BSTA_513_26S/>

-   Complete the [when2meet](https://whenisgood.net/fpgdi2q) for office hours

    -   If your calendar feels set, take 5 minutes to fill this out now!

    -   **Complete by Thursday at 11pm!!!**

-   Highly suggest that you make an appointment with a learning specialist through [Student Academic Support Services](https://sakai.ohsu.edu/portal/site/Student_Support)!

```{r}
#| echo: false
countdown::countdown(5)
```

## Syllabus

-   No changes from last quarter! 

## Key Course Info (at the top of the syllabus)

-   If an assignment on Sakai is closed or you are submitting late work, **please email me AND the TAs your work** (see [instructor page](instructors.qmd) for all email addresses)
-   If you are emailing for an extension, please email me AND the TAs. (see [instructor page](instructors.qmd) for all email addresses)
-   Latework policy
    -   **For homework:** can be turned in any time before course ends, but you will not receive feedback on late work
    -   **For labs:** one "no questions asked", 3-day extension
        -   After that, 1 point deducted per day late
-   **Attendance policy:** attend in person and fill out exit tickets for 12 out of the 19 classes
    -   There are **FIVE required in-person attendance days** (4 lab discussions and 1 project poster day)
-   The class will end on `{r} w11d1 %>% format("%B %d, %Y")`. **All coursework MUST be completed by `{r} last_day %>% format("%B %d, %Y")`**
-   Feel free to fill out the [Ongoing Course Feedback](https://forms.office.com/Pages/ResponsePage.aspx?id=V3lz4rj6fk2U9pvWr59xWFMopmPUjRtDiHLswhgacNhURDc2S1dGNTlTRVJVUFoyQUUzNFJMS0JXUi4u) with any concerns or questions.

## Assessment Breakdown

-   We will have quizzes and required in-person lab discussions (on schedule)

|  |  |  |  |  |
|:--------------|:--------------|:--------------|:--------------|:--------------|
| **Course activity** | **Type of Assessment** | **Due Date** | **Percentage of final grade (BSTA 513)** | **Percentage of final grade (BSTA 613)** |
| Homework | Formative | Every 1-2 weeks | 30% | 25% |
| Quizzes | Summative | `r q1_close %>% format("%m/%d")`, `r q2_close %>% format("%m/%d")`, `r q3_close %>% format("%m/%d")` | 25% | 25% |
| Project Labs | Formative/Summative | Every 2-3 weeks | 25% | 25% |
| Lab Discussion + Poster Day Participation | N/A | Every 2-3 weeks | 5% | 5% |
| Project Poster | Summative | `r proj %>% format("%m/%d")` | 10% | 10% |
| Exit tickets (Attendance) | N/A | Twice Weekly | 5% | 5% |
| 613 Readings | Formative | Approx. every other week | 0% | 5% |


## How to print slides

-   Anyone have issues with this?

    -   I can show how to do it in Chrome and Safari

-   [Instructions on Quarto page](https://quarto.org/docs/presentations/revealjs/presenting.html#print-to-pdf)

