---
title: "Lab 3: Feedback and Discussion"
author: "Nicky Wakim"
title-slide-attributes:
    data-background-color: "#C2352F"
date: "`r library(here); source(here('class_dates.R')); w7d1`"
format: 
  revealjs:
    theme: "../simple_NW.scss"
    chalkboard: true
    slide-number: true
    show-slide-number: all
    width: 1955
    height: 1100
    footer: "Lab 3: Feedback and Discussion"
    html-math-method: mathjax
    highlight-style: ayu
execute:
  echo: true
  freeze: auto  # re-render only when source changes
---

```{r}
#| echo: false
#| warning: false
#| message: false

library(tidyverse)
library(gt)
library(janitor)
library(rstatix)
library(knitr)
library(gtsummary)
library(moderndive)
library(broom) 
library(here) 
library(ggplot2)
library(ggpubr)
library(boot) # for inv.logit()
library(lmtest)
library(epiDisplay)

```

## Lab 3 Feedback

- I'll have more feedback on Wednesday after the TAs grade and I meet with them


## Get into groups of 2-4

- No more than 4! 
- Introduce yourself and your research question
- Share you html documents with each other (email, airdrop, etc.)

```{r}
#| echo: false
countdown::countdown(5)
```

## Any issues with your forest plots?

- See if you can help each other troubleshoot
- Some minor changes may not be worth it, but it's good to make things easier for yourself later

```{r}
#| echo: false
countdown::countdown(10)
```

## Discussion of research questions and results

- Going back to your research question, how would the table, forest plot, and interpretations fit into your results?
- Look at the [poster template](https://ohsuitg-my.sharepoint.com/:p:/r/personal/wakim_ohsu_edu/Documents/Teaching/Classes/BSTA_513_26S/Student_files_BSTA_513/Project/Resources/poster_template.pptx?d=wb113c78c33e4497fab393030df6448e6&csf=1&web=1&e=TPyCn9)
- OR start looking at Labs 1-3: what can you start placing in your poster?

```{r}
#| echo: false
countdown::countdown(15)
```

## Questions to consider

- How did you handle any categorical variables?
  - Income, family size, number of children, etc
- Are the categories for the forest plot in order?
- How did you interpret your estimated odds ratios?
  - If you have many categories in your main variable, do you see a trend? 
  - What might be a better way to report a trend?
- If you had interactions, how would your interpretation change?
